---
title: "Milwaukee Bucks Analysis"
font: 12pt
date: "April 1"
author: "Alex Matz"
output:
  html_document: default
  word_document: default
  pdf_document: default
---


```{r Libraries, message=FALSE, warning=FALSE, include=FALSE}
library(devtools)
library(nbastatR) ## Library which includes many types of data from the NBA. This requires installation from gitHub
library(remotes) ##Package which is necessary to get court data from package rNBA
library(rNBA) ##Package with court dimension data directly from NBA site
library(tidyverse)
library(gridExtra)
library(gganimate)
library(gifski)
```

```{r Data Read in, warning=FALSE, message=FALSE, include=FALSE}
Sys.setenv("Vroom_Connection_Size" = 131072*2) ## Sets the vroom allocation so data can be pulled in

Bucks_2000<-teams_shots(teams = "Milwaukee Bucks",  ## This Loads in the Milwaukee Bucks Shot Data for 2000
                   seasons= 2000,
                   team_ids = NULL, all_active_teams = F, 
                   season_types = "Regular Season",
                   measures = "FGA",
                   periods = 0,
                   months = 0,
                   date_from = NA,
                   date_to = NA,
                   nest_data = F,
                   return_message = T
                   )

Bucks_2005<-teams_shots(teams = "Milwaukee Bucks", ## This Loads in the Milwaukee Bucks Shot Data for 2005
                   seasons= 2005,
                   team_ids = NULL, all_active_teams = F, 
                   season_types = "Regular Season",
                   measures = "FGA",
                   periods = 0,
                   months = 0,
                   date_from = NA,
                   date_to = NA,
                   nest_data = F,
                   return_message = T
                   )

Bucks_2010<-teams_shots(teams = "Milwaukee Bucks", ## This Loads in the Milwaukee Bucks Shot Data for 2010
                   seasons= 2010,
                   team_ids = NULL, all_active_teams = F, 
                   season_types = "Regular Season",
                   measures = "FGA",
                   periods = 0,
                   months = 0,
                   date_from = NA,
                   date_to = NA,
                   nest_data = F,
                   return_message = T
                   )
Bucks_2015<-teams_shots(teams = "Milwaukee Bucks", ## This Loads in the Milwaukee Bucks Shot Data for 2015     
                   seasons= 2015,
                   team_ids = NULL, all_active_teams = F, 
                   season_types = "Regular Season",
                   measures = "FGA",
                   periods = 0,
                   months = 0,
                   date_from = NA,
                   date_to = NA,
                   nest_data = F,
                   return_message = T
                   )
Bucks_2020<-teams_shots(teams = "Milwaukee Bucks",  ## This Loads in the Milwaukee Bucks Shot Data for 2020
                   seasons= 2020,
                   team_ids = NULL, all_active_teams = F, 
                   season_types = "Regular Season",
                   measures = "FGA",
                   periods = 0,
                   months = 0,
                   date_from = NA,
                   date_to = NA,
                   nest_data = F,
                   return_message = T
                   )

```
```{r Data Merging and Explore, message=FALSE, warning=FALSE, include=FALSE}
Bucks_All<- rbind(Bucks_2000,Bucks_2005, Bucks_2010, Bucks_2015, Bucks_2020) ## This combines all the data into one dataframes.

##head(Bucks_All)

## locationX & locationY are the shot location variables
## isShotAttempted is logical variable which is always true
## isShotMade is a logical boolean of true and false
```

```{r Court Makeup, message=FALSE, warning=FALSE, include=FALSE}
court_coorinates<- courtOutline(full=T) ## From the rNBA package, this obtains the coordinates of the court

##court_coorinates<-court_coorinates%>% ## Groups by the court parts (Type)
 ##  group_by(type)

##summary(court_coorinates$type) ##Gives the parts of the court listed

court <- ggplot()+geom_line(data=court_coorinates, aes(x=x,y=y, group=type)) ## This makes the court given the rNBA coordinates

```

```{r Plotting, echo=FALSE, message=FALSE, warning=FALSE}

### Plot is shot location on the created court plot ###

animated_shots<- court+geom_point(data = Bucks_All, aes(x=locationX, y=locationY, color=typeEvent, alpha=.01))+ ##Adds the shot location points on the court
   scale_color_manual(values = c("Made Shot"="olivedrab", "Missed Shot"= "grey"))+ ## Colors the shots for made or missed
   guides(alpha="none")+ ##removes other guides
   labs(color="Shots", x="", y="")+ ##Changes the labels of the legend and the axis
   transition_time(yearSeason)+ ##creates transition animations
  labs(title = 'Season: {frame_time}') ##Adds changing title to animation


animate(animated_shots, fps=.25,nframes = 5) ## Runs Animation

##Goal is to see the the shot density on the court for every season.
```


```{r Find the NBA Shot Distances and Percentages, echo=FALSE, message=FALSE, warning=FALSE}

shot_distance<- Bucks_All%>% ##Makes the mean shot distant variable
   group_by(yearSeason)%>%
   summarise(mean(distanceShot))

Bucks_All<-left_join(Bucks_All, shot_distance, by="yearSeason") ##Join the new mean shot distance variable to the main data set across all years


shot_percentage<- Bucks_All%>%
   select(yearSeason,isShotMade)%>% ##Selects the variables for the new variable
   group_by(yearSeason)%>%
   summarise(mean(isShotMade)*100)##Calculates the the total shot percentage using the mean of the boolean is shot made

Bucks_All<-left_join(Bucks_All, shot_percentage, by="yearSeason") ##Join the new shot percentage/year to the main data set across all years


Bucks_All<-rename(Bucks_All, "ave_shot_dist"="mean(distanceShot)") ## Rename
Bucks_All<-rename(Bucks_All, "shot_percent"="mean(isShotMade) * 100") ## Rename



Shot_Percentage_Plot<- ggplot()+geom_line(data=Bucks_All,aes(x=yearSeason, y=shot_percent)) +labs(x="Season", y="Average Shooting Percentage")+  ## Labels
   theme_classic() ## Basic Theme to remove grid lines

Shot_Distance_Plot<-ggplot()+geom_line(data=Bucks_All,aes(x=yearSeason, y=ave_shot_dist))+
   labs(x="", y="Average Shot Distance (ft)", title = "Trends of Shooting over 20 Years")+ ## Labels
   theme_classic() ## Basic Theme to remove grid lines

grid.arrange(Shot_Distance_Plot, Shot_Percentage_Plot) ## Plots

## The goal here is to see trends of shot distance and see if shooting percentage goes with it.


```

```{r Heights Compared to Distance, message=FALSE, warning=FALSE, include=FALSE}
Players_Distance<-Bucks_All%>% ##Makes Shot Distance for each Player per season
   group_by(idPlayer, yearSeason)%>%
   summarise(mean(distanceShot))
 
Players_Distance<-rename(Players_Distance, "player_shoot_distance"="mean(distanceShot)") ## Rename  

Players_Distance<- Players_Distance %>% ##Select only the 2 variables for the merge
   select(player_shoot_distance)
   
Bucks_All<-left_join(Players_Distance,Bucks_All, by="idPlayer") ## Joins Shot Distance into main data set


Player_Profiles<-player_profiles(player_ids = Bucks_All$idPlayer) ## Attains player profile data from NBAstatR

Bucks_All<-inner_join(Bucks_All,Player_Profiles, by="idPlayer") ##Joins the Player Profiles Df with main df

height_shot_distance<- ggplot(data = Bucks_All, aes(x=heightInches, y=player_shoot_distance))+geom_point()+facet_wrap(~yearSeason)+labs(x="Player Heights (in)", y="Average Shot Distance", title="Players Average Shot Distance by their Heights")
```

```{r Height and Distance Plot, echo=FALSE, message=FALSE, warning=FALSE}
height_shot_distance

###The goal here was to see if the players heights effected the amount of shots taken at greater distances. These chunks are seperated solely supress warnings
```

Main Points

* Over the years the density of the shots attempted appear to grow in distance from the hoop.
* A indication of change of game play is evident with 2020 being an example with more shots being behind the 3 point line then all the other seasons.
* Even as shot distance grows, the shooting percentage goes up in the later seasons. This may indicate that training improved shot percentage which allowed shots to be made at greater distance. 
*  Most of these shots taken from greater distances are the shorter players. They is an apparent trend that in later seasons, taller players take shots at greater distances but still not as many as the shorter players.



Further Reading

This analysis was done on the seasons 1999-2000, 2004-2005, 2009-2010, 2014-2015, 2019-2020 for solely on the team Milwaukee Bucks. The data was source from two R packages, rNBA and NBAstatR. These packages were designed to retrieve data from valid sources including NBA.com. Using coordinate shot data compared against the court coordinates, using plotting it was found that there is a trend in the shot distance increasing over the 5 seasons. These trends became clearer when animating the seasons and the shot locations. Deeper analysis showed that although the shot distance increase, the shooting percentage did not decrease overall. Instead they appear to follow the same trend indicating that training may have increased shooting percentage therefore allow players to shoot farther from the basket. 

Looking at what type of players most often shot at these greater distances, average shooting distances were plotted against there heights for the seasons. The Bucks have had a variety of players during these seasons but many have been players over the 20 year period so the heights were often the same across seasons. It was found that for all seasons, the shorter players shot consistently at greater distances. This may because the positions of the players. In 2020, taller players shot farther which may suggest another trend change. 

Overall, over the 5 seasons, shot distance increased along with the shooting percentage. Most of these shots were taken by shorter players but there may be new trend of taller players also shooting farther.
  
  
  


